---
title: "Introduction"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(rAIutils)
```

My plan for this rAIutils package is to collect various functions for working with generative AI that exist in Python but not (yet) in R. These functions will likely be written mostly by state-of-the-art LLMs with guidance and editing by me.

rAIutils is currently not on CRAN. However, you can install it from GitHub with the GitHub-install package of your choice, such as `remotes::install_github("smach/rAIutils")` or `pak::pak("smach/rAIutils")`. 

For now, there's just one function in the package: `parse_pdf_to_markdown()`, which is an R wrapper around using LlamaCloud's LLamaParse API to turn a PDF file into markdown. While there are other ways to generate text from PDF files that don't require a cloud service and API key, I've been impressed by LlamaParse. 

LlamaParse is a paid service, but there's a generous free tier of 10,000 credits per month. The standard parser uses 3 credits per page (there's a faster one that's for text only and not tables), which means more than 3,300 pages per month -- enough for me. The major drawback to LlamaParse is if you are working with sensitive data you want to keep private (I've been using it for PDFs already published on the public Web, so not an issue).

See details on how to get a free LlamaCloud API key at [https://docs.cloud.llamaindex.ai/llamacloud/getting_started/api_key](https://docs.cloud.llamaindex.ai/llamacloud/getting_started/api_key).

`parse_pdf_to_markdown()` can handle PDFs that are local files on your system or at public URLs on the Web. It requires two arguments: input source (a character string with the location of the file) and api_key.

Sample code for parsing a file, if you've saved your API key in an environment variable called `LLAMA_PARSE_API_KEY`:

```
library(rAIutils)
my_api_key <- Sys.getenv("LLAMA_PARSE_API_KEY")
my_file  <- "https://cran.r-project.org/web/packages/RColorBrewer/RColorBrewer.pdf"
markdown_from_file <- parse_pdf_to_markdown(input_source = my_file, api_key = my_api_key)
```

Note that the request may take awhile to process. The function checks every 8 seconds to see if the job has finished.

If you want to save results to a file, you can run something like:

`cat(markdown_from_file, file = "RColorBrewer.md")`

